---
title: "ETC1010: Introduction to Data Analysis"
week: "Week 10, part A"
subtitle: "Regression and Decision Trees"
author: "Nicholas Tierney & Stuart Lee"
email: "nicholas.tierney@monash.edu"
date: "May 2020"
pdflink: ""
bgimg: "images/bg1.jpg"
output:
  xaringan::moon_reader:
    css:
      - ninjutsu 
      - "assets/animate.css"
      - "assets/monash-logo.css"
      - "assets/monash-brand.css"
      - "assets/monash-fonts.css"
      - "assets/styles.css" # small improvements
      - "assets/custom.css" # add your own CSS here!
      - "assets/demo.css" # this should be removed
    self_contained: false 
    seal: false 
    chakra: 'libs/remark-latest.min.js'
    lib_dir: libs
    includes:
      in_header: "assets/custom.html"
    mathjax: "assets/mathjax-local/MathJax.js?config=TeX-AMS-MML_HTMLorMML"
    nature:
      highlightStyle: github 
      highlightLanguage: r 
      highlightLines: true
      highlightSpans: false 
      countIncrementalSlides: false
      slideNumberFormat: '%current%/%total%'
      navigation:
        scroll: false 
        touch: true
        click: false
      ratio: '16:9'
---
  
```{r setup, include=FALSE}
library(tidyverse)
library(tibble)
library(vctrs)
library(knitr)
library(kableExtra)
library(countdown)
library(knitr)
library(lubridate)
library(gridExtra)
library(plotly)
library(broom)
library(broomstick)

opts_chunk$set(echo = TRUE,   
               message = FALSE,
               warning = FALSE,
               collapse = TRUE,
               fig.height = 4,
               fig.width = 8,
               fig.retina = 2,
               out.width = "100%",
               fig.align = "center",
               cache = FALSE)
```

```{r funs, echo = FALSE}
draw_split <- function(split){

df_add_sst <- df %>% 
  mutate(x_split = if_else(x <= split, "left", "right")) %>% 
  group_by(x_split) %>% 
  mutate(y_bar = mean(y),
         y_slope = 0) 

ggplot(df_add_sst, 
       aes(x = x, 
           y = y)) + 
  geom_point() +
  geom_vline(xintercept = split, linetype = 2) +
  geom_line(aes(y = y_bar), colour = "salmon") + 
  geom_segment(mapping = aes(xend = x, 
                             yend = y_bar), 
               color = "grey50",
               alpha = 0.5) +
  scale_x_continuous(breaks = seq(-0.5, 0.5, 0.1))
}
```

  
```{r titleslide, child="components/titleslide.Rmd"}
```

---
# recap

- Decision Trees

---
# Overview

- What is a regression tree?
- How is it computed?
- Deciding when its a good fit
  - rmse
- Comparison with linear models
- Using multiple variables
- Next class:
  - How a classification tree differs from a regression tree?



When the response is categorical, the model is called a classification tree. The criteria for making the splits changes also. There are a number of split criteria commonly used. If we consider a binary response ($y=0, 1$), and $p$ is the proportion of observations in class $1$.

- Gini: $2p(1-p)$
- Entropy: $-p(\log_e p)-(1-p)\log_e(1-p)$

Which rewards splits where the observations are all one class.