---
title: "ida-exercise-5b-extension"
date: "21/04/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Cricket data

Using the information from this vignette here: https://docs.ropensci.org/cricketdata/

## Use the cricketdata package to extract the T20 batting data of male players for your two favorite countries.

```{r}
library(cricketdata)

```


## Between 2010 and 2015, which country has the better record? Explain the reason for your answer, and your analysis choices.

```{r}

```



# Extension Example: Crossrates 

An example we have seen is the cross rates data available at [https://openexchangerates.org/](https://openexchangerates.org/). To access this data you need to:

1. Get a free plan from https://openexchangerates.org/signup/free
2. Tell this function your API key -- `Sys.setenv("OER_KEY" = "your-key-here")`

Then you can access the data using a command like:

```
u <- sprintf(
    "https://openexchangerates.org/api/historical/%s.json?app_id=%s",
    day, Sys.getenv("OER_KEY")
  )
res <- jsonlite::fromJSON(u)
```

There's a nice help page here: https://gist.github.com/njtierney/0fe2daa11292e046f83fec7d8e9a8164

## Pull the crossrates for the month of August 2018

```{r}

```

## Make a plot to compare the crossrates for AUD, EUR, GBP, NZD, JPY. Make a descriptive summary of what you learn.