---
title: "Assessment"
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE} 
library(knitr)
library(tidyverse)
library(kableExtra)

opts_chunk$set(echo = FALSE)

```


```{r create-assessment}

assessment_df <- data.frame(
  "assessment_task" = c("Reading Quizzes",
                        "Lab Exercises",
                        "Assignment 1",
                        # "Mid-semester in-class exam",
                        "Assignment 2",
                        "Data Analysis Practical Exam",
                        "Project",
                        "Final Exam"),
  "value" = c("5%",
              "5%",
              "6%",
              # "8%",
              "6%",
              "15%",
              "13%",
              "50%"),
  "due_date" = c("Weeks 1-8",
                 "Each class period",
                 "Week 4 (8th April)",
                 # "Week 6 (20th April)",
                 "Week 8 (12th May)",
                 "Week 11 (27th May)",
                 "Week 12 (8th June)",
                 "TBA")
  )

names(assessment_df) <- c("Assessment Task",
                          "Value",
                          "Due Date")

assessment_df %>% 
kable(escape = FALSE, 
      align = c("l", "c", "l")) %>% 
  kable_styling(bootstrap_options = "striped",
                fixed_thead = T) %>% 
  column_spec(1, width = "0.5cm") %>% 
  column_spec(2, width = "1cm") %>% 
  column_spec(3, width = "3cm") 
```

## Assignments

Information to complete the assignments is below:

- Unzip the file and double click on the rstudio project, "assignment-1-2020-s1.Rproj" This contains the data and Rmd file
- **Note that the assignment is also on rstudio.cloud as well, named after the assessment task. So Assignment 1 will be named, "Assignment 1".**


```{r assignments}
assignment_df <- tibble(
  Assignment = 1,
  `Due Date` = "2020/04/01 at 5pm",
  Instructions = "[instructions](https://ida.numbat.space/assignments/assignment-1-instructions)",
  Zipfile = "[zipfile](https://ida.numbat.space/assignments/ida-assignment-1.zip)"
)

assignment_df %>% 
kable(escape = FALSE, 
      align = c("l", "c", "l", "l")) %>% 
  kable_styling(bootstrap_options = "striped",
                fixed_thead = T) %>% 
  column_spec(1, width = "0.5cm") %>% 
  column_spec(2, width = "1cm") %>% 
  column_spec(3, width = "3cm") 
```

## Project

[Project instructions](project/project-instructions.html) 

- **Deadline 1 (6th May)**:  Find team members, and potential topics to study (List of groups will be posted here)
- **Deadline 2 (13th May)**:  Team members and team name, data description.
- **Deadline 3 (20th May) **:  Electronic copy of your data, and a page of data description, and cleaning done, or needing to be done.
- **Deadline 4 (29th May) **:  Final version of story board uploaded. 
- **Presentation in week 12**

There are several ways to do the storyboard:
  - `learnr` package that the lecture notes are written in. [Example files with css](project/lecture10b.zip)
  - [flexdashboard](https://rmarkdown.rstudio.com/flexdashboard/) package, and there is an [example here](project/pisa_au_app.zip)
	- or a basic [shiny](https://shiny.rstudio.com/gallery/) app
  
Past Example projects from ETC1010:

- [Analysing Movie Data](https://ebsmonash.shinyapps.io/Thor/)
- [Instagram analysis](https://ebsmonash.shinyapps.io/Instagram/)

- Examples from different web sites:
  - [Blue Bikes Sharing in Boston](https://www.r-bloggers.com/blue-bikes-sharing-in-boston/)
  - [What birds are observed near Radolfzell? Bird occurrence data in R](https://ropensci.org/blog/2018/08/21/birds-radolfzell/)
	- [Who wrote the anti-Trump New York Times op-ed? Using tidytext to find document similarity](https://www.r-bloggers.com/who-wrote-the-anti-trump-new-york-times-op-ed-using-tidytext-to-find-document-similarity/)
	- [Hotels vs Airbnb - Barcelona case study (proof of concept)](https://nycdatascience.com/blog/student-works/web-scraping/hotels-vs-airbnb-barcelona-case-study-proof-of-concept/)
	- [Don't Worry: Google Only Checks Your Location Every 10 Minutes](http://www.sastibe.de/2018/04/don-t-worry-google-location/)
	- [World Income, Inequality and Murder](http://staff.math.su.se/hoehle/blog/2018/07/09/gini.html)
	- [Text analysis of Trump's tweets confirms he writes only the (angrier) Android half](http://varianceexplained.org/r/trump-tweets/)
	- [What digits should you bet on in Super Bowl squares?](http://varianceexplained.org/r/super-bowl-squares/)
	- [Exploring handwritten digit classification: a tidy analysis of the MNIST dataset](http://varianceexplained.org/r/digit-eda/)
	- [Facebook advertising case-study: dog food and manhood](https://rickpackblog.wordpress.com/2018/08/20/facebook-advertising-case-study-dog-food-and-manhood/)
	- [Analysis of Los Angeles Crime with R](https://datascienceplus.com/analysis-of-los-angeles-crime-with-r/)
	- [Looking at Seinfeld dialogues using #tidytext](https://pradeepadhokshaja.wordpress.com/2018/08/06/looking-at-seinfeld-dialogues-using-tidytext/)
	- [Twitter Coverage of the useR Conference 2018](https://github.com/neilfws/Twitter/blob/master/user2018/code/R/user2018.md)

## Exam information

- [Midterm 2019](exam/midterm2019.pdf) [(solution)](exam/midterm2019_solution.pdf)

- [Formula sheet](exam/formula_sheet.pdf)

- [Final Exam for 2018 Semester 1](exam/exam2018.pdf) [(solution)](exam/exam2018_solution.pdf)
- [Final Exam for 2018 Semester 2](exam/exam2018_S2.pdf) [(solution)](exam/exam2018_S2_solution.pdf)
- [Final Exam for 2019 Semester 1](exam/exam2019-s1.pdf)