---
title: "ETC1010: Introduction to Data Analysis"
subtitle: "Semester 1 2020"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
library(tidyverse)
library(knitr)
library(kableExtra)
library(glue)
library(googlesheets4)
library(lubridate)
library(ymlthis)
library(yaml)


monash_pals <- c(blue = "#4269a5",
                 purple = "#8e4f89", 
                 green = "#39825a", 
                 orange = "#dc6141")
```

```{r functions}
generate_readings <- function(week, part){
  readings <- read_yaml("materials/readings.yml")
  reading_list <- glue('readings[["Week {week}"]]\\
                                      [["Readings"]]\\
                                      [["{part}"]]')
  eval(parse(text = reading_list)) %>% 
    glue_collapse("\n\n")
}

```

- Lecturer: Dr. Nicholas Tierney
- Chief Examiner: Di Cook
- Tutors: Steph Kobakian, Sherry Zhang, Cynthia, Sarah Belet, Nitika Kandhari

This is the course website. This page shows information about the consultation times, MoVE, and Schedule and lecture notes.

The "Assessments" tab above provides information about all assessment for ETC1010.

## Expectations 

- Unless your question is a personal issue or an administrative matter, please do not send us direct email. The [ED site](https://edstem.org/) is appropriate for questions related to the course material.
- For personal or administrative issues, the email contact is: ETC1010.Clayton-x@monash.edu
- ED is used for asking for help, reading and lab exercise quizzes and assignment turn in. Only your name and email is stored in this system.
- **If you register after the start of the semester** or if you
  miss a lecture/lab, it is your responsibility to catch up with
  missed material, learn about due dates for material to be turned in,
  and getting assigned to a group for team work.

## Consultation times (Consults from Week 2)

- Steph: TBA
- Sarah: Monday 3-4pm
- Sherry: Friday 1pm-2pm
- Nitika: Tuesday 11:30am-12:30pm
- Cynthia: Thursday 12pm-1pm
- Nick: Monday morning **Menzies W1105**
- Nick: Wednesday morning **Menzies W1105**

## Tentative Schedule

There are two lectorials every week:

* **Monday 4-6pm** (G31 - LTB Bldg 92)
* **Wednesday 12 - 2pm** (G31 - LTB Bldg 92)

There are no lectorials during the midsemester break.

```{r read-create-sheet, eval = FALSE}
# to be run interatively when google sheet is updated
google_schedule <- read_sheet("https://docs.google.com/spreadsheets/d/1OmwIYBtuedFICwqXToTNx1xRRoDO0ZjjVjumvmg-TYU/edit#gid=0")

write_csv(google_schedule, path = "materials/schedule.csv")
```


```{r read-data-google}
schedule <-  read_csv("materials/schedule.csv") %>% 
  mutate(Slides = case_when(
    !is.na(HTML) & !is.na(PDF) ~ glue("<a class='fa fa-link' href='slides/{HTML}'></a> <a class='fa fa-file-pdf' href='slides/{PDF}'></a>"),
    !is.na(HTML) ~ glue("<a class='fa fa-link' href='slides/{HTML}'></a>"),
    !is.na(PDF) ~ glue("<a class='fa fa-file-pdf' href='slides/{PDF}'></a>"),
    TRUE ~ ""
  ),
  Exercise = case_when(
    !is.na(Exercise) ~ glue("<a class='fa fa-link' href='{Exercise}'></a>"),
    TRUE ~ ""
  ),
  Topic = ifelse(is.na(Topic), "", Topic),
  Exercise = ifelse(is.na(Exercise), "", Exercise)) %>% 
  # Date = janitor::excel_numeric_to_date(Date)) %>% 
  select(Week, 
         Lecture, 
         Date, 
         Slides, 
         Topic, 
         Exercise, 
         Readings, 
         Assessment) %>% 
  # generate readings
  mutate(Readings = map2_chr(Week, Lecture, generate_readings)) 

midsem_break <- 10 # after which week (but there are 2 lectures per week)

schedule %>% 
  kable(escape = FALSE, 
        align = c("c", "c", "c", "c", "l", "l")) %>% 
  kable_styling(bootstrap_options = "striped",
                fixed_thead = T) %>% 
  column_spec(1, width = "0.5cm") %>% 
  column_spec(2, width = "1cm") %>% 
  column_spec(3, width = "3cm") %>% 
  column_spec(4, width = "1cm") %>% 
  group_rows(group_label = "Midsemester Break", 
             start_row = midsem_break + 1, 
             end_row = midsem_break, 
             label_row_css = "background-color: #d9e1e8;") %>% 
  add_indent(1:midsem_break) %>%  # this needs to be done to match up with above
  add_indent((midsem_break+1):24) # this needs to be done to match up with above
```



