# Demonstrate with koala bilby data (write this as live code)

Here is a little data set to practice `pivot_longer`, `pivot_wider` and `separate` on.

First we load the library `tidyverse` to load up all of our functions.

```{r library}
library(tidyverse)
```


```{r koala-bilby}
kb <- read_csv(here::here("exercises/2a/data/koala_bilby.csv"))
kb
```

---

### Exercise 1

Pivot the data into long form, naming the two new variables, `label` and `count`

```{r kb-long}
kb_long <- kb %>% 
  pivot_longer(cols = -ID,
               names_to = "label", 
               values_to = "count")

kb_long
```

### Exercise 2

Separate the labels into two new variables, `animal`, `state`

```{r kb-separate}
kb_long_sep <- kb_long %>%
  separate(col = label, 
           # into = c(__, ___))
           into = c("animal", "state"))

kb_long_sep
```

### Exercise 3

Pivot the long form data into wide form, where the columns are `state`s. 

```{r kb-wide-state}
kb_long_sep %>% 
  # pivot_wider(names_from = ___, 
  pivot_wider(names_from = state,
              values_from = count)
```

### Exercise 4

Convert the long form data into wide form, where the columns are the `animal`s. 

```{r kb-pivot-wider-animal}
kb_long_sep %>% 
  # pivot_wider(key = ___, 
  pivot_wider(names_from = animal,
              values_from = count)
```
