---
title: "ETC1010: Introduction to Data Analysis"
week: "Week 10, part A"
subtitle: "Week of introduction"
author: "Nicholas Tierney"
email: "nicholas.tierney@monash.edu"
date: "11th Mar 2020"
pdflink: ""
bgimg: "images/bg1.jpg"
output:
  xaringan::moon_reader:
    css:
      - ninjutsu 
      - "assets/animate.css"
      - "assets/monash-logo.css"
      - "assets/monash-brand.css"
      - "assets/monash-fonts.css"
      - "assets/styles.css" # small improvements
      - "assets/custom.css" # add your own CSS here!
      - "assets/demo.css" # this should be removed
    self_contained: false 
    seal: false 
    chakra: 'libs/remark-latest.min.js'
    lib_dir: libs
    includes:
      in_header: "assets/custom.html"
    mathjax: "assets/mathjax-local/MathJax.js?config=TeX-AMS-MML_HTMLorMML"
    nature:
      highlightStyle: github 
      highlightLanguage: r 
      highlightLines: true
      highlightSpans: false 
      countIncrementalSlides: false
      slideNumberFormat: '%current%/%total%'
      navigation:
        scroll: false 
        touch: true
        click: false
      ratio: '16:9'
---
  
```{r titleslide, child="components/titleslide.Rmd"}
```


```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(emo)
library(ggraph)
library(igraph)
library(tidygraph)
library(knitr)
library(lubridate)
library(gridExtra)
library(plotly)
library(countdown)
opts_chunk$set(echo = TRUE,   
               message = FALSE,
               warning = FALSE,
               collapse = TRUE,
               fig.height = 4,
               fig.width = 8,
               out.width = "100%",
               fig.align = "center",
               cache = FALSE)

as_table <- function(...) knitr::kable(..., format='html', digits = 3)
```


---

# Announcements

- Current alternative for viewing: Download the HTML or the RMD for the moment and view locally.
- I'm in the process of sorting it out

---
# The Practical Exam

- Time: 1 Hour
- Task: Complete a data analysis and answer questions (like the assignments, but short - similar to the in class exercises)
- Conditions: 
  - Open book
  - Internet access
  - No contacting / messaging / communicating friends during exam (penalty will be the same as cheating in exam)

---
# The Project Data Milestone: What do Submit

- An Rmd file + 
- The rendered HTML file
- The data
- Zipped / compressed together
- In the document:
  - Write about the data - what is it, what are the variables, what sort of questions are you asking, what did you need to do clean the data, what cleaning is remaining?
  - (About 1-2 paragraphs of text.)

---
# The Project: More general questions

- "How many questions should we ask?"
  - About 3 - roughly one per person in your group
- "How will the project be marked?"
  - You will all mark the projects as you watch them
- You will need to present for 5 minutes max - use this to help frame how long your analysis should be. 
- 5 minutes is short - but it is actually a pretty serious challenge!
- But also consider that the project will be marked by me as well - you won't need to cover everything

---
# The project: What should we present?

Many options!

- You can submit an Rmarkdown document + HTML file
- You present with a [flexdashboard](https://rmarkdown.rstudio.com/flexdashboard/) (demo)
- You could even present as a slide show (like the lecture slides)
- How to share - Ideally you can use netlify drag and drop (demo)

---
# Exam details

- Worth 50% of your final grade
- NOT Open Book
- Calculator: Yes - Allowed to take in "Only HP 10bII+ or Casio FX82 (any suffix) calculator permitted"
- One A4 page of notes, printed on both sides. This will be collected at the end of the examination.
- **Hurdle requirement**: You must get 40% on the exam to pass the course (also true of the final project).

---
# Exam Details

- I'll now talk about the questions in the exam, and some of the concepts you need to be familiar with
- These concepts will help guide what you focus on in the lectures
- Disclaimer: This list is not exhaustive - these are to help give you a sense of what I'm thinking about for each of the questions in the exam.

---
# Tidy data Concepts:

- Variables
- Observations
- Tidy data

---
# Data Wrangling Concepts:

- Converting "messy" data to tidy data
- Code / key functions to use to convert data into "tidy" data
  - e.g, `pivot_longer`, `pivot_wider`, `separate`, etc.

---
# Relational data concepts:

- Left Join
- Why do joins of data
- Predict output of a join
- Sketching out code to summarise data from a join

---
# Data visualisation concepts:

- How the grammar of graphics produces a plot - identify plots produced by which code
- Understand the focus of a given graphic on the data - questions like: 
  - "What and how does this graphic make us focus on what feature of the data?"
  - "What do you learn from a graphic?"

---
# Missing Values concepts:

- Principles of tidy missing data
- Interpreting graphics of missing data
- Predict output of function on data with missing values
- Imputation - what is it, which methods are good / bad / better

---
# Linear Models Concepts:

- Write down an equation of a model from code output

--

- The formula $y = 3x + 5$ is a function with input $x$, and output $y$, when x is `___` , the output is `___`

---
# Linear Models Concepts:

$$\widehat{height_{in}} \sim 3.62 + 0.78 Width_{in}$$
- **slope**: For each additional inch the painting is wider, the height is expected to be higher, on average, by 0.78 inches.
- **Intercept**: Paintings that are 0 inches wide are expected to be 3.62 inches high, on average.
- If a paining is 5 Inches wide, what is it's estimated height?

---
# Linear Models Concepts:

- How to make predictions from a fitted model
- Understand what makes predictions good and bad
- Measurements of model fit:
  - R2 - what it is - what values mean good/bad fit?
  - **A good idea to bring in an equation of R2**
- Residuals - What do we expect to see
- Centering variables
- Think about how you can improve fit of models to your data

---
# Programming concepts:

- Why write functions
- How to write a function in R
- How to take existing code and turn it into a function
- Identify potential mistakes in provided code
- Understand what `map` does

---
# Networks Concepts:

- From an association / correlation matrix, which are most or least related?
- Understand how you can convert a numeric matrix into a binary association matrix
- Understand how a correlation matrix (or other association) of data relates to a provided network (let's do an example)

---
# How to study for the exam

These concepts guide what to focus on

While the readings aren't assessed, **they provide great information that will certainly help improve your understanding**

Remember that these are to help you focus, I can't give you the exam, but I can help tell you what is important.

Disclaimer: This list is not exhaustive - these are to help give you a sense of what I'm thinking about for each of the questions in the exam.

---
# Learning technique: study

- Practice previous exam in exam conditions (last semester's exam is on dmac.netlify.com)

--

- Think up exam questions, write them down

--

- Practice explaining concepts - either out aloud by yourself, in the mirror, to a friend, to an empty room. Saying things out aloud builds better connections.

---
# Learning technique: in exam

--

- Peruse the exam before starting

--

- Take off one shoe (My high school teacher claimed this worked)

--

- Number questions in order from easy-hard - start with the easiest one first

--

- Work out a marks : minutes ratio. (e.g.,60 marks with 60 minutes means one mark per minute)


---


---
# On a more personal note

I've had a great time this semester

I've been **seriously impressed** with you all - you are all so bright!

Thanks for being engaged and curious, I'm really excited to see your projects!

--

Also a special thanks to everyone for dealing with the 8am lecture!

---

```{r endslide, child="components/endslide.Rmd"}
```
