---
title: "ETC1010: Introduction to Data Analysis"
week: "Week 6, part A"
subtitle: "Style, file paths, & functions"
author: "Nicholas Tierney"
email: "ETC1010.Clayton-x@monash.edu"
date: "April 2020"
pdflink: ""
bgimg: "images/bg1.jpg"
output:
  xaringan::moon_reader:
    css:
      - ninjutsu 
      - "assets/animate.css"
      - "assets/monash-logo.css"
      - "assets/monash-brand.css"
      - "assets/monash-fonts.css"
      - "assets/styles.css" # small improvements
      - "assets/custom.css" # add your own CSS here!
      - "assets/demo.css" # this should be removed
    self_contained: false 
    seal: false 
    chakra: 'libs/remark-latest.min.js'
    lib_dir: libs
    includes:
      in_header: "assets/custom.html"
    mathjax: "assets/mathjax-local/MathJax.js?config=TeX-AMS-MML_HTMLorMML"
    nature:
      highlightStyle: github 
      highlightLanguage: r 
      highlightLines: true
      highlightSpans: false 
      countIncrementalSlides: false
      slideNumberFormat: '%current%/%total%'
      navigation:
        scroll: false 
        touch: true
        click: false
      ratio: '16:9'
---
  
```{r titleslide, child="components/titleslide.Rmd"}
```


```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
opts_chunk$set(echo = TRUE,   
               message = FALSE,
               warning = FALSE,
               collapse = TRUE,
               fig.height = 4,
               fig.width = 8,
               fig.align = "center",
               cache = FALSE)

as_table <- function(...) knitr::kable(..., format='html', digits = 3)
```


# Recap

- Missing Data
- Web Scraping

---
# Upcoming due dates

- Assignment 2: 13th May (Released this week)
- Practical Exam: 3rd June
- Project: 8th October (See examples of past projects in assessments)

---
# Practical Exam?

- A live data analysis
- 1 Hour 

---
# Project?

- Collect / find your own data
- Clean the data
- Determine interesting questions to answer about the data
- Plan how to execute analysis of the data
- Communicate the idea, data cleaning, and analysis (oral presentation)
- Further details are on the course website

---
# Lecture Overview

- Organising your own folders
- File paths and Rstudio projects
- (Intro to) Using functions

---
# File Paths and organising yourself

- It's important when you start working on your own machine that you understand _file storage hygiene_. 
- It helps prevent unexpected problems and makes you more productive 
- You'll spend less time fighting against strange file paths. 
- Not sure what a file path is? We will explain that as well!

---
class: transition
# Your Turn: Breakout groups

1. What your normal "workflow" is for starting a new project / assessment
2. Possible challenges that might arise when maintaining your project / assessment
3. What is a file path?

---
# What even is a file path?
  
- This all might be a bit confusing if you don't know what a file path is.
- A file path is: "the machine-readable directions to where files on your computer live." 
- So, this file path:

```
/Users/njtierney/rmd4sci-materials/demo-gapminder.Rmd
```

Describes the location of the file "demo-gapminder.Rmd". 


---
# What even is a file path

We could visualise this path:

```
/Users/njtierney/rmd4sci-materials/demo-gapminder.Rmd
```

as:

```
users
  └── njtierney
        └──  rmd4sci-materials
               └── demo-gapminder.Rmd
```

---
# What even is a file path

- To read in the `gapminder.csv` file, you might need to write code like this:

```r
gapminder <- read_csv("/Users/njtierney/Desktop/rmd4sci-materials/data/gapminder.csv")
```
 
This is a problem, because this is not portable code.

---
# A Mantra: Start a new project \- start an RStudio project

- This section is heavily influenced by [Jenny Bryan's great blog post on project based workflows.](https://www.tidyverse.org/articles/2017/12/workflow-vs-script/)
- Sometimes this is the first line of an R Script or R markdown file.

```r
setwd("c:/really/long/file/path/to/this/directory)
```
- What do you think the `setwd` code does?

---
# What does `setwd()` do?

- "set my working directory to this specific working directory".

- It means that you can read in data and other things like this:

```r
data <- read_csv("data/mydata.csv")
```

- Instead of

```r
data <- read_csv("c:/really/long/file/path/to/this/directory/data/mydata.csv")
```

---
# Using `setwd()`

- This has the effect of **making the file paths work in your file**
- This is a problem because, among other things, using `setwd()`:
  -  Has 0% chance of working on someone else's machine (**this includes you in >6 months**)
  - Your file is not self-contained and portable. (Think: _"What if this folder moved to /Downloads, or onto another machine?"_)
- To get this to work, you need to hand edit the file path to your machine. 
- This is painful. And when you do this all the time, it gets old, fast.

--

If you have an RStudio project file inside the `rmd4sci-materials` folder, you can instead write the following:

```r
gapminder <- read_csv("data/gapminder.csv")
```

---
class: transition

# Your Turn: Whole class discussion

- (1-2 minutes) What folders are above the `health.csv` file in the following given file path? 

`"/Users/miles/etc1010/week1/data/health.csv"`

- and the result of using the below code in `demo-gapminder.Rmd`, then using the code, and then moving this to another location, say inside your C drive?

```r
setwd("Downloads/etc1010/week1/week1.Rmd)
```

---
# Is there an answer to the madness?

- This file path situation is a real pain. 
- Is there an answer to the madness?

--

The answer is yes!

I highly recommend when you start on a new idea, new research project, paper. Anything that is new. It should start its life as an **rstudio project**. 

---
# Rstudio projects

An rstudio project helps keep related work together in the same place. Amongst other things, they:

* Keep all your files together
* Set the working directory to the project directory
* Starts a new session of R
* Restore previously edited files into the editor tabs
* Restore other rstudio settings
* Allow for multiple R projects open at the same time.

---
# Rstudio projects

This helps keep you sane, because:

* Your projects are each independent. 
* You can work on different projects at the same time. 
* Objects and functions you create and run from project idea won't impact one another.  
* You can refer to your data and other projects in a consistent way.

And finally, the big one

**RStudio projects help resolve file path problems**, because they automatically set the working directory to the location of the rstudio project.

---
# The "here" package

- RStudio projects help resolve file path problems
- In some cases you might have many folders in your r project. To help navigate them appropriately, you can use the `here` package to provide the full path directory, in a compact way.

```r
here::here("data")
```

returns

```
[1] "/Users/njtierney/Desktop/rmd4sci-materials/data"
```

---
# The `here` package


```r
here::here("data", "gapminder.csv")
```

returns

```
[1] "/Users/njtierney/Desktop/rmd4sci-materials/data/gapminder.csv"
```

You can read the above `here` code as:

> In the folder `data`, there is a file called `gapminder.csv`, can you please give me the full path to that file? 

---
# The `here` package


This is really handy for a few reasons:

1. It makes things _completely_ portable
1. Rmarkdown documents have a special way of looking for files, this helps eliminate file path pain.
1. If you decide to not use RStudio projects, you have code that will work on _any machine_

---
class: transition
# Remember

> If the first line of your R script is 

```
setwd("C:\Users\jenny\path\that\only\I\have")
```
> I will come into your office and SET YOUR COMPUTER ON FIRE 🔥.

-- Jenny Bryan

---
# Aside: How to create an RStudio project

- Go to [section 5.12 of rmarkdown for scientists](https://rmd4sci.njtierney.com/workflow.html#aside-creating-an-rstudio-project)

---
# Summary of file paths and rstudio projects


In this lesson we've:

- Learnt what file paths are
- How to setup an rstudio project
- How to construct full file paths with the `here` package

---
class: transition 
# Recommendations on how to file structure in ETC1010

---
# File structures for class

Approach 1:  Folder per week

```
/Users/njtierney/etc1010/week_1/
```

```
users
  └── njtierney
      └──  etc1010
            └── etc1010.Rproj
            └── week_1
                └── lecture_1.html
                └── lecture_1.pdf
                └── ida-exercise-1.Rmd
                └── data
                    └── file.csv
            └── week_2
                └── lecture_2.html
                └── lecture_2.pdf
                └── ida-exercise-2.Rmd
                └── data
                    └── file.csv
                
```

---
# File structures for class

Approach 2: flater structure

```
/Users/njtierney/etc1010/
```

```
users
  └── njtierney
      └──  etc1010
            └── etc1010.Rproj
            └── lecture_1.html
            └── lecture_1.pdf
            └── ida-exercise-1.Rmd
            └── data
                └── data.csv
```

---
class: transition

# Remember: There is no one true "correct" file format

It's just important to have a system


---
class: motivator

# Motivating Functions

---
# Do you see any problems with this code?

```{r eval=FALSE}
st_episode <- st %>%
  html_nodes(".np_right_arrow .bp_sub_heading") %>%
  html_text() %>%
  str_replace(" episodes", "") %>%
  as.numeric()

got_episode <- got %>%
  html_nodes(".np_right_arrow .bp_sub_heading") %>%
  html_text() %>%
  str_replace(" episodes", "") %>%
  as.numeric()

twd_episode <- got %>%
  html_nodes(".np_right_arrow .bp_sub_heading") %>%
  html_text() %>%
  str_replace(" episodes", "") %>%
  as.numeric()
```

---
# Next Lecture: Why functions?

- Automate common tasks in a power powerful and general way than copy-and-pasting:
    - You can give a function an evocative name that makes your code easier to understand.
    - As requirements change, you only need to update code in one place, instead of many.
    - You eliminate the chance of making incidental mistakes when you copy and paste (i.e. updating a variable name in one place, but not in another).

--

- Down the line: Improve your reach as a data scientist by writing functions (and packages!) that others use

---

# Take the lab quiz!
