---
title: "Introduction to functions"
author: "YOUR NAME"
output: html_document
---

```{r setup}
library(tidyverse)
library(polite)
library(rvest)
```

# Take some input, return some output

Functions are often described as "take in some input, return some output"

While that can be useful, I am not entirely conviced that is what makes them awesome and useful, but we can discuss a few examples here.

## We've already used functions

- mean
- median
- min
- max

```{r}
x <- 1:10
x
mean(x)
median(x)
min(x)
max(x)
```

They take in some input, and then return some output.

But say that we wanted to calculate the difference between the minimum and maximum, we could do:

```{r}
max(x) - min(x)
```

But we can give this a more descriptive name, and turn it into a function like:

```{r}
range_diff <- function(x){
  max(x) - min(x)
}
```

This then takes the same input, and gives us some output:

```{r}
range_diff(x)
```

So that's one way to write a function. 

# Web scraping example

Now let's discuss an example from class, which we'll go into more detail in class next week.

# Scrape the list of most popular TV shows on IMDB: http://www.imdb.com/chart/tvmeter

```{r}
url <- "http://www.imdb.com/chart/tvmeter"
check_url <- bow(url)

tv_data <- scrape(check_url)

scrape_show_info <- function(x){
  title <- x %>%
    html_node("#title-overview-widget h1") %>%
    html_text() %>%
    str_trim()
  runtime <- x %>%
    html_node("time") %>%
    html_text() %>%
    str_replace("\\n", "") %>%
    str_trim()
  genres <- x %>%
    html_nodes(".txt-block~ .canwrap a") %>%
    html_text() %>%
    str_trim() %>%
    paste(collapse = ", ")
  tibble(title = title, runtime = runtime, genres = genres)
}
```


```{r}
st %>%
  html_nodes("#title-overview-widget .subtext") %>%
  html_text() %>%
  str_trim() %>%
  str_remove_all("\\n") %>%
  str_split("\\|") %>%
  map(str_trim)

```


```{r}
tv_html <- bow("http://www.imdb.com/chart/tvmeter") %>% scrape()

library(glue)

web_stubs <- tv_html %>%
  html_nodes(".titleColumn a") %>%
  html_attr("href") %>%
  paste("http://www.imdb.com", ., sep = "")
```

```{r}
scrape_show_info(web_stubs[1])
web_stubs[2]

web_stubs[1]
top_show_check_1 <- bow(web_stubs[1])
top_show_data_1 <- scrape(top_show_check)
scrape_show_info(top_show_data_1)

top_show_check_2 <- bow(web_stubs[2])
top_show_data_2 <- scrape(top_show_check_2)
scrape_show_info(top_show_data_2)

top_show_check_3 <- bow(web_stubs[3])
top_show_data_3 <- scrape(top_show_check_3)
scrape_show_info(top_show_data_3)

bow_scrape_show_info <- function(x){
  bow(x) %>% scrape() %>% scrape_show_info()
}

bow_scrape_show_info(web_stubs[1])
bow_scrape_show_info(web_stubs[2])
bow_scrape_show_info(web_stubs[3])
```

# Why `()` and `[]`

`()` are used for functions - so `mean(x)` runs the function `mean` on `x`.
`[]` are used for indexing. so `x[1]` gets the first thing from x.

```{r}
x <- 1:10
x[1]
x[2]
x[1:2]
```

 

```{r}


title <- 
  x %>%
    html_node("#title-overview-widget h1") %>%
    html_text() %>%
    str_trim()
  runtime <- x %>%
    html_node("time") %>%
    html_text() %>%
    str_replace("\\n", "") %>%
    str_trim()
  genres <- x %>%
    html_nodes(".txt-block~ .canwrap a") %>%
    html_text() %>%
    str_trim() %>%
    paste(collapse = ", ")
  tibble(title = title, runtime = runtime, genres = genres)
```



## Examine each of the *first three* (or however many you can get through) tv show subpage to also obtain genre and runtime.


## Time permitting, also try to get the following:

  - How many episodes so far
  - First five plot keywords
  - Country
  - Language
  
Add this information to the data frame you created in step 1.


