---
output: html_document
---

```{r setup}
library(tidyverse)
library(rpart)
library(rpart.plot)
```

# Classification Trees

Here we will practice fitting classification trees, for a categorical
variable that has more than two classes

## Measurements of Australian Crabs

This data consists of measurements about crab species in WA (described 
[here](http://ggobi.org/book/chap-data.pdf))

> Measurements on rock crabs of the genus Leptograpsus. One
species L. variegatus had been split into two new species, previously grouped
by color, orange and blue. Preserved specimens lose their color, so it was
hoped that morphological differences would enable museum specimens to be
classified. There are 50 specimens of each sex of each species, collected on site
at Fremantle, Western Australia. 


```{r}
crabs <- read_csv("data/australian-crabs.csv")
glimpse(crabs)
```

What each variable does? 

 * __species__: orange or blue
 * __sex__: male or female
 * __index__: observation identifier 1–200
 * __frontal lip (FL)__ length, in mm
 * __rear width (RW)__ width, in mm
 * __carapace length (CL)__ length of midline of the carapace, in mm
 * __carapace width (CW)__ maximum width of carapace, in mm
 * __body depth (BD)__ depth of the body; for females, measured after
displacement of the abdomen, in mm

## Aim

Can we determine the species and sex of the crabs based on the five
body variables above? 

## Your turn: create a predictor variable 

Make a new predictor called `class` that combines the `species` and `sex`
variables:

```{r}
crabs <- crabs %>% 
  mutate(class = ---(species, sex)) %>% 
  mutate(class = factor(class))

crabs %>% 
  count(class)
```


Make plots of each explantory variable (FL:BD) coloured by the class

## Your turn: create a decision tree

```{r}
rp_crab_all <- rpart(class ~ ., data = select(crabs, class, FL:BD))

rpart.plot(rp_crab_all)
```

Which variables appear to be important? How many terminal nodes are there?

## Plot the decision boundaries for the two most important variables

```{r}
crab_grid <- ## use expand grid?
crab_grid$sex <- predict(rp_crab_all, newdata=crab_grid, type="class")
ggplot(data=crab_grid, aes(x=FL, y=RW, color=class)) + 
  geom_point(alpha=0.3, size=3) + 
  geom_point(data=crabs, aes(x=FL, y=RW, color=class), shape=2, size=3) + 
  scale_colour_brewer(palette="Dark2") +
  theme_bw() + 
  theme(aspect.ratio=1, legend.position="none") 

```


## Your turn: modify the input parameters

* try changing the splitting criterion (from 'gini' to 'importance')
* minsplit and minbucket

How do your results compare to the default? Trying plotting the decision
boundaries again and look at the metrics for your model fit.

