---
output: html_document
---

```{r setup}
library(tidyverse)
library(rpart)
library(rpart.plot)
```

## Manually computing splits

```{r}
exams <- read_csv("data/exam.csv")
exams
```

```{r}
ggplot(exams, aes(x=Exam1, y=Exam2, color=factor(Label))) +
  geom_point(size=4) + 
  scale_color_brewer("", palette="Dark2") +
  theme_minimal() +
  theme(text = element_text(size=20)) 
```

First, we are going to work out the number of misclassifications



```{r}
# given a binary label in the data, what is the most frequent class 0 or 1?
max_vote <- function(data, y_var) {
    max(
    c(
    sum(data[[y_var]] == 0),
    sum(data[[y_var]] == 1)
    )
  )
}

# our 'criterion' for determining a split
# total = the number of cases to be classified,
# left the left partition,
# right the right partition,
# y_var the class label
miss_class <- function(total, left, right, y_var) {
  l_max <- max_vote(left, y_var)
  r_max <- max_vote(right, y_var)
  total - (l_max + r_max)
}

miss_party <- function(sorted_data, x_var, y_var) {
  # total number of cases
  total <- nrow(sorted_data)
  # initialise a data frame to hold the results
  miss_class_ans <- data.frame(
    x = numeric(total-1), 
    n = integer(total-1)
  )
  colnames(miss_class_ans) <- c(x_var, "n_miss")
  # loop over the data to generate the number of classifications
  for(i in 2:total){
    left <- sorted_data[1:(i-1), ]
    right <- sorted_data[i:total, ]
    miss_class_ans[i-1,][[x_var]] <- mean(sorted_data[c(i-1, i),][[x_var]])
    miss_class_ans[i-1,][["n_miss"]] <- miss_class(total, left, right, y_var)
  }
  miss_class_ans
}

```

Now we are ready to count the miss-classifications for the first 
variable `Exam1`. Remember in order to create the partitions we need to 
sort the data by the variable we are trying to split:

```{r}
exam1_misclass <- exams %>% 
  arrange(Exam1) %>% 
  miss_party(., "Exam1", "Label")
```
Plot the result: 

```{r}
ggplot(exam1_misclass, aes(x = Exam1, y = n_miss)) +
  geom_line()
```


Which value of `Exam1` minimises the the number of misclassifications and
should be used for the split?


Try again for exam 2:

```{r}
exam2_misclass <- exams
  arrange(Exam2) %>% 
  miss_party(., "Exam2", "Label")

ggplot(exam2_misclass, aes(x = Exam2, y = n_miss)) +
  geom_line()
```

What is the minimum number of misclassifications? 
Which value(s) of `Exam2` minimises the number of misclassifications? There
are actually two!

So the split happens on the first variable `Exam1` because the split
results in fewer mis-classifcations!

Run the `rpart` function and plot the tree:

```{r}
rpart(factor(Label) ~ ., data = exams)
```

Try running the miss party function after filtering to values on Exam1 below
the first split: why do you think `rpart` didn't make a further split at this
node?

```{r}
exams
  filter(Exam1 < 56.74626) %>% 
  arrange(Exam2) %>% 
  miss_party(., "Exam2", "Label")
```
