---
author: ""
title: ""
output: html_document
---


```{r setup}
library(tidyverse)
library(rwalkr)
library(lubridate)
```

# What to do if you are stuck?

* Read through the lecture slides again, they most likely have an answer to these questions!
* Take the question and copy-paste it into google, adding "lubridate" to the google query.

# Play with dates

- Convert something of the format "2020-01-20" into a date?

```{r add-tz-ymd}
___("2020-01-20")
```

- How do you show today's date?
```{r show-today}
___()
```

Convert something of the format "2020-01-20 10:00:01" into a date time?

```{r ymd-hms}
***_***("2020-01-20 10:00:01")
```

Get the weekday out of "2020-01-20"

```{r wday-and-options}
___("2020-01-20")
```

Get the month number out of "2020-01-20"

```{r month}
month("2020-01-20")
```

Get the month name out of "2020-01-20"

```{r month-label}
month("2020-01-20", label = TRUE)
```

get the unabbreviated month name out of "2020-01-20"

```{r month-label-abbr}
___("2020-01-20", ___ = TRUE, ___ = FALSE)
```

Get the day of the year out of "2020-01-20"

```{r yday}
___("2020-01-20")
```

# Go back to lecture slides

# Next part of the exercise

```{r rwalkr-load}
walk_all <- melb_walk_fast(year = 2019)

walk <- walk_all %>% filter(Sensor == "Melbourne Central")

# write_csv(walk, path = "exercises/3b/data/walk_2019.csv")
# write_csv(walk, path = "data/walk_2019.csv")
```

```{r read-walk}
walk <- read_csv("data/walk_2019.csv")
# walk <- read_csv("exercises/3b/data/walk_2019.csv")
walk
```


# Lab exercise

## Pull the 2019 data for the sensor Birrarung Marr. Make plots to answer these questions. 

```{r walk-burra}
walk_birra <- walk_all %>% filter(Sensor == "Birrarung Marr")
```

## Is there the same daily pattern as observed at Melbourne Central? If not, how does it differ?

> Hint: combine the Birrarung Marr data with Melbourne central data (walk) using `bind_rows()`. Or even subset the original data that has all the obsercations into something that has melbourne central AND birrarung marr. There are many ways to get to the same answer.

```{r bind-both}
walk_both <- bind_rows(walk_birra,
                       walk)
```

## Can you detect special events, like the Australian Open tennis, based on pedestrian counts?  

> hint: try using a calendar plot from the lecture slides

```{r calendar-plot}

```



## Can you match these with known events, found from web searching?

# Extension exercises (to be covered at a later time)

## Can you summarise this into weekly information?
##