---
title: "Exercise 7B Part Two"
author: "YOUR NAME"
output: html_document
---

```{r pkgs}
library(tidyverse)
library(readxl)
library(janitor)
```


## Your turn: new data - fertility data

- Using the fertility data from the [gapminder web site](https://www.gapminder.org/data/) (in data/ folder)
- This contains data of number of babies born per woman
- You can find this out by exploring the data in the aforementioned gapminder website (https://www.gapminder.org/data/documentation/gd008/)
-  Conduct the same analysis as in the gapminder example from the lecture ( ida-exercise-7b.Rmd), but for life expectancy.
- Find unusual countries

The code below will help you read in the data and process it, but you will need to make some changes to do the full analysis.

```{r read-data}
fertility_raw <- read_xlsx(
  "data-raw/indicator-undata-total-fertility.xlsx"
  ) %>%
  clean_names() %>%
  rename(country = total_fertility_rate)

fertility_raw
```

You will need to pivot the data into longer form, using the `pivot_longer` function. 

You should know how to do the rest from your experience with past assignments.

```{r tidy-fertility}
fertility <-
  fertility_raw %>%
  pivot_longer(cols = ___, # everything but country,
               # what is the name of the new variable we are changing the
               # column of names to?
               names_to = ___,
               # What is the name of the column we are changing the values to?
               values_to = ___) %>%
  extract out the year information
  mutate(year = ___) %>%
  # filter so we only look at years above 1950
  filter( ___) %>%
  # center year around 1950
  mutate(year1950 = ___)
```

plot the fertility data for each year

```{r eval=FALSE}
ggplot(fertility, 
       aes(x = year, 
           y = babies_per_woman, 
           group = country)) +
  geom_line(alpha=0.1)
```

Now try and answer these questions below. You can use the examples in `ida-exercise-7b.Rmd` to help you answer these questions.

# How has fertility changed over years? For each country?

## How has fertility changed in Australia?

## Fit a model for Australia

## Explore model fit for Australia

## How has fertility increased every year / decade in Australia?

## How well does the model fit?

## Fit a linear model for all countries

## Plot the model fit against the data in two plots

## plot the model coefficients with plotly

## Summarise what you have learned from the model coefficients

## Find those countries with the worst fit

