---
# (if time) Example: AFL Finals tweets

The `rtweet` package allows you to pull tweets from the archive. It gives only the last 6-9 days worth of data. You need to have a twitter account, and you need to create an app (its really basic) in order to pull twitter data. The instructions that come from this package (https://rtweet.info) are pretty simple to follow.

---
# (if time) Example: AFL Finals tweets


Given that it is AFL final week, I thought it might be interesting to look at tweets that use the hashtag "#AFLFinals". Once you have a developer account, this is as simple as 

```
afl <- search_tweets(
  "#AFLFinals", n = 20000, include_rts = FALSE
)
```

---
# (if time) Example: AFL Finals tweets

Here is the data collected in the previous year's AFL finals.

```{r load-afl}
afl <- read_rds("data/afl_twitter_past.rds")
afl
```

---
# Your turn

- When was the final played last year?
- What is the range of dates of this data?
- Who is the most frequent tweeter using this hashtag?
- Are there some days that have more tweets than others?
- Are there some hours of the day that are more common tweet times?
]

---
# Your Turn: Sentiment analysis

We need to break text of each tweet into words, tag words with sentiments, and make a cumulative score for each tweet.

- Which tweeter is the most positive? negative?
- Is there a day that spirits were higher in the tweets? Or when tweets were more negative?
- Does the tweeter `aflratings` have a trend in positivity or negativity?


---
  
  
  
# Your Turn: Sentiment analysis

```{r}
load("data/afl_twitter.rda")
afl
```

We need to break the text of each tweet into words, tag the words with sentiments, and make a cumulative score for each tweet.

- Which tweeter is the most positive? negative?
- Is there a day that spirits were higher in the tweets? Or when tweets were more negative?
- Does the tweeter `aflratings` have a trend in positivity or negativity?

```{r eval=FALSE, echo=FALSE}
afl_sentiment <- afl %>% 
  # group by something
  group_by() %>%
  # unnest the tokens
  unnest_tokens() %>%
  inner_join(get_sentiments("afinn")) %>%
  summarise(sentiment = mean(score, na.rm=T)) 

afl_sentiment <- afl %>% 
  select(status_id, 
         screen_name, 
         created_at, 
         text) %>% 
  left_join(afl_sentiment, by=c(???))

afl_sentiment %>% 
  group_by(???) %>%
  ???(s = mean(sentiment, na.rm=TRUE)) %>% arrange(desc(s))
afl_sentiment %>% mutate(day = ???)) %>% ggplot(aes(x=day, y=sentiment)) + geom_point() + geom_smooth(se=FALSE)
afl_sentiment %>% filter(screen_name == ???) %>% 
  mutate(day = ???) %>%
  ggplot(aes(x=day, y=sentiment)) + geom_point() + geom_smooth()
```


# Your turn

- When was the final played last year?
- What is the range of dates of this data?
- Who is the most frequent tweeter using this hashtag?
- Are there some days that have more tweets than others?
- Are there some hours of the day that are more common tweet times?


# Sentiment analysis

We need to break the text of each tweet into words, tag the words with sentiments, and make a cumulative score for each tweet.

