---
output: html_document
---

# Exploring sentiment in Jane Austen

`janeaustenr` package contains the full texts, ready for analysis for for Jane Austen's 6 completed novels: 

1. "Sense and Sensibility"
2. "Pride and Prejudice"
3. "Mansfield Park"
4. "Emma"
5. "Northanger Abbey"
6. "Persuasion"


---
# Exploring sentiment in Jane Austen

```{r show-jane-austen}
library(janeaustenr)
library(stringr)

tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, 
                                     regex("^chapter [\\divxlc]", 
                                           ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)
```

# Count joyful words in "Emma"

```{r count-joy}
nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")

tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)
```

# Count joyful words in "Emma"

"Good" is the most common joyful word, followed by "young", "friend", "hope". 

All make sense until you see "found". 

Is "found" a joyful word?

# Questions: 

- What are the most common "anger" words used in Emma?

```{r}
nrc_anger <- get_sentiments("nrc") %>% 
  filter(sentiment == "anger")

tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_anger) %>%
  count(word, sort = TRUE)
```

```{r}
common_emotion <- function(text_df, book_name, emotion){
  nrc_emotion <- get_sentiments("nrc") %>% 
  filter(sentiment == emotion)

text_df %>%
  filter(book == book_name) %>%
  inner_join(nrc_emotion) %>%
  count(word, sort = TRUE)
}
```

```{r}
common_emotion(tidy_books, "Emma", "surprise")
common_emotion(tidy_books, "Pride & Prejudice", "joy")
```


- What are the most common "surprise" words used in Emma?

# Next round of questions:

- Using the bing and afinn lexicon compute the proportion of positive words in each of Austen's books.

```{r}
austen_sentiment_bing <- tidy_books %>%
  inner_join(get_sentiments("bing")) 

austen_sentiment_bing %>% 
  group_by(book) %>% 
  count(___) %>% 
  mutate(prop_total = n / sum(n))
```


```{r}
austen_sentiment_afinn <- tidy_books %>%
  inner_join(get_sentiments("afinn"))
```

- Which book is the most positive? negative?