---
title: "Lecture 7A Exercises"
author: "your name"
output: html_document
---
```{r setup}
library(tidyverse)
library(lubridate)
library(broom)
```

## Let's fit a model to CO2

The Scripps CO2 program records atmospheric carbon dioxide measurements
from sampled since 1957 at various stations around the world along the same
longitude (why?) This link 
https://scrippsco2.ucsd.edu/data/atmospheric_co2/sampling_stations.html
shows the different stations and provides links to the data. 

We have extracted and downloaded data for the South Pole. 
If you have a look at this csv file, at what line does the data start?
```{r data-read}
co2_spo <- read_csv(
  # here::here("exercises/7a/data/daily_merge_co2_spo.csv"), 
  here::here("data/daily_merge_co2_spo.csv"),
                    col_names = c("date", 
                                  "time", 
                                  "day", 
                                  "decdate", 
                                  "n", 
                                  "flg", 
                                  "co2"), 
                    skip = 69) %>%
  mutate(lat = -90.0, 
         lon = 0, 
         stn = "spo") %>%
  filter(flg == 0) %>% # the flag recognises whether the value is accepted
  mutate(date = ymd(date)) %>%
#   First we re-scale to start from 1 until the number of days in the time frame. 
  mutate(day0 = day - min(day))

co2_spo
```

### Task: plot date against co2 with points

```{r gg-co2-spo}
ggplot(co2_spo, 
       aes(x = date, y = co2)) + 
  geom_point()
```

### Task: fit a linear model, using `day` variable as the explanatory variable. 

And now tidy it up

```{r lm-day}
co2_day_fit <- lm(co2 ~ day, data = co2_spo)
tidy(---)
```

# calculate the summary of day0 and day to compare `day` against `day0`

```{r lm-summary}
summary(co2_spo$day)
summary(co2_spo$day0)
```

# Question: "What is range of dates of this data?"

- The variables, day and day0 range from ...

## Task: fit a linear model, predicting co2 using day0

```{r fit-co2}
co2_fit <- lm(co2 ~ day0, data = co2_spo)
co2_fit
```

Now we tidy up the output and store information on the intercept and slope so we can describe the fit

```{r tidy-up}
library(broom)
tidy_co2_fit <- tidy(---)
tidy_co2_fit
int <- round(tidy_co2_fit$estimate[1], 4)
slope <- round(tidy_co2_fit$estimate[2], 4)
```

Then the fitted model will be:

$$\widehat{co2}=$$ `r int` + `r slope`.

# Task: Predict from the model

- For day0 = 10000 
- the model predicts co2 to be `r int`+`r slope`*10000 = `r int + slope * 10000`. 

## Question: What date is day0=10000? (Hint: use filter and select on the data to work it out)"


# Plot the model

We can use the `augment` function to extract residuals and fitted values, and pointwise diagnostics

```{r augmented-model2}
co2_model <- augment(co2_fit, co2_spo)

co2_model
```

plot the date and co2 with points and add in the line

```{r augmented-model}
ggplot(co2_model, 
       aes(x = date, 
           y = co2)) + 
  geom_point() + 
  geom_line(aes(y = .fitted), colour = "blue")
```

### Examine residuals

- **Residuals** are calculated for each observed $y$ by computing the difference: $y_i-\widehat{y_i}$.

- Plotting the residuals against fitted values (or x in simple linear models) can reveal problems with the fit.

```{r residuals-date}
ggplot(co2_model, 
       aes(x = date, 
           y = .resid)) + 
  geom_point() 
```

# Assessment of fit

- The relationship between co2 and day0 is nonlinear!
- The linear model does a reasonable job of explaining the increasing trend over time, but it especially mismatches the observed data at the ends, and middle of the time period.

### Your turn add a quadtratic term

- Try to add a quadratic term (in day0), or more, to the model to improve the fit. You may want to centre the day0 values, or even standardise them, to get a nice quadratic form. 
- Repeat a similar process to what you have seen earlier when exploring model fit

```{r daysq-mod }
co2_sq <- co2_spo %>%
  mutate(day0sq = (day0-median(day0))^2)

co2_fit2 <- lm(co2 ~ day0 + day0sq, data = co2_sq)

tidy(---)
```

## Plot the model fit 

```{r daysq-mod}
co2_model2 <- augment(---, co2_sq)

ggplot(co2_model2, 
       aes(x = date, y = co2)) + 
  geom_point() + 
  geom_line(aes(y = ---), colour="blue")
```


## Plot the residuals

```{r daysq-resid}
ggplot(co2_model2, 
       aes(x = .fitted, y = .resid)) + 
  geom_point()
```


## Your Turn: Predict co2 at another location

Using your model, built using values collected the south pole sensor, see how well it fits values from Point Barrow, Alaska. 

1. Download the data. You can use almost the same code as for SPO but check the file name at https://scrippsco2.ucsd.edu/data/atmospheric_co2/ptb. See here:

https://scrippsco2.ucsd.edu/assets/data/atmospheric/stations/merged_in_situ_and_flask/daily/daily_merge_co2_ptb.csv

```{r read-ptb}
co2_ptb <- read_csv(
  file = ---,
                    col_names = c("date", 
                                  "time", 
                                  "day", 
                                  "decdate", 
                                  "n",
                                  "flg", 
                                  "co2"), 
                    skip = 69) %>%
  mutate(lat = 71.3, 
         lon = 156.6, 
         stn = "ptb") %>%
  filter(flg == 0) %>%
  mutate(date = ymd(date)) %>% 
  mutate(day0 = day - min(co2_spo$day))
```


2. The code below is a way to fit new data to the previously fit `co2_fit2`. You need to ensure you have the same variables as the data used in `co2_fit2` (the data used to fit that was called `co2_sq`)

```{r augment}
co2_ptb_sq <- co2_ptb %>% 
  mutate(co2_sq = ___)
co2_model_ptb <- augment(co2_fit2, newdata = co2_ptb)
```

3. Plot the data, and overlay the fitted model. You can use code like this.

```{r}
ggplot(co2_model_ptb, aes(x=date, y=co2)) + 
  geom_point() + 
  geom_line(aes(y=.fitted), colour="blue")
```


## Extension exercise


### Seasonality

- If you plot the residuals on a short time frame, you can see that there is some seasonality. Values are high in spring, and low in autumn! 
- Brainstorm with your table members - yes, please talk with each other - ideas on how to fit a model that takes seasonality into account. There are multiple solutions, and maybe some that we haven't thought of.

```{r eval=FALSE, echo=FALSE}
ggplot(filter(co2_model2, year(date)>2002, year(date)<2005), aes(x=date, y=.std.resid)) + 
  geom_point()
```
