---
title: "Sentiment"
author: "Your Name"
date: "10/05/2020"
output: html_document
---

## Sentiment

- One way to analyze the sentiment of a text is to consider the text as a combination of its individual words 

- and the sentiment content of the whole text as the sum of the sentiment content of the individual words

- essentially a dictionary where different words are categorised either as
positive or negative or on a numeric scale

```{r}
afinn <- get_sentiments("afinn") # numeric
bing <- get_sentiments("bing") # categorical
```


After tokenising into words, use a left/inner join to get the words sentiments.

Here's what the sentiment "This will be an uncertain time for us my love"

```{r}

text <- c("This will be an uncertain time for us my love",
          "I can hear the echo of your voice in my head",
          "Singing my love",
          "I can see your face there in my hands my love",
          "I have been blessed by your grace and care my love",
          "Singing my love")

text
text_df <- tibble(line = seq_along(text), text = text)

text_df
```


```{r}
text_df %>% 
  # filter(line == 1) %>%
  unnest_tokens(word, text) %>% 
  left_join(afinn) %>% 
  group_by(line) %>%
  summarise(avg_senti = mean(value, na.rm = TRUE))
```

```{r}
text_df %>% 
  # filter(line == 1) %>%
  unnest_tokens(word, text) %>% 
  left_join(bing) %>% 
  count(sentiment)
```

## Your turn

Using the quotes we looked at above use the "afinn" lexicon to compute the
average sentiment of each quote. Which one is considered more positive?


