---
output: html_document
---

# Lab exercise

- OECD PISA, what factors affect reading scores?
- 15 year old standardised test scores, Australia, 2015
- Response: math
- Predictors: gender, anxtest, wealth, math_time, books, tvs

- Make a plot of all the variables

```{r fig.width=12, fig.height=6, eval=FALSE, echo=FALSE}
load("data/pisa_au_sub.rda")
pisa_au <- pisa_au %>% 
  filter(!is.na(gender)) %>%
  filter(!is.na(anxtest)) %>%
  filter(!is.na(wealth)) %>%
  filter(!is.na(???)) %>%
  filter(!is.na(books)) %>%
  filter(!is.na(???))
p1 <- ggplot(pisa_au, aes(x=???, y=math)) +
  geom_boxplot() 
p2 <- ggplot(pisa_au, aes(x=anxtest, y=math)) +
  geom_point() + geom_smooth(se=FALSE)
p3 <- ggplot(pisa_au, aes(x=???, y=math)) +
  geom_point() + geom_smooth(se=FALSE)
p4 <- ggplot(pisa_au, aes(x=math_time, y=math)) +
  geom_point() + geom_smooth(se=FALSE)
p5 <- ggplot(pisa_au, aes(x=factor(???), y=math)) +
  geom_boxplot() 
p6 <- ggplot(pisa_au, aes(x=factor(tvs), y=math)) +
  geom_boxplot() 
grid.arrange(p1, p2, p3, p4, p5, p6, ncol=3)
```

---
# Fit a linear model

```{r eval=FALSE, echo=FALSE}
pisa_lm <- lm(math~gender+???+wealth+math_time+books+???,
              data=pisa_au, weights=stuweight)
summary(pisa_lm)
```

---
# Fit a regression tree

```{r eval=FALSE, echo=FALSE}
pisa_rp <- rpart(math~gender+???+wealth+math_time+books+???, data=pisa_au, weights=stuweight)
pisa_rp
```

```{r eval=FALSE, echo=FALSE}
rpart.plot(???)
```

---
# What is the most important variable

```{r eval=FALSE, echo=FALSE}
ggplot(pisa_au, aes(x=???, y=math)) +
  geom_point() + geom_vline(xintercept=???, colour="hotpink")
```

---
# How good is the model? Compute the $R^2$ for the tree.

```{r eval=FALSE, echo=FALSE}
pisa_rp_aug <- cbind(pisa_au, e=???(pisa_rp))
1-sum(pisa_rp_aug$e^2)/sum((pisa_au$math-???(pisa_au$math))^2)
```

---
# Which model fits better? The tree or the linear regression model?

- Change the control parameters to reduce the $R^2$ of the tree below that of the regression model.

```{r eval=FALSE, echo=FALSE}
pisa_rp <- rpart(math~gender+???+wealth+math_time+???+tvs, data=pisa_au, weights=stuweight, control = rpart.control(???))
pisa_rp
pisa_rp_aug <- cbind(pisa_au, e=???(pisa_rp))
1-sum(pisa_rp_aug$e^2)/sum((pisa_au$math-???(pisa_au$math))^2)
```

