---
title: "Harry Potter"
author: "Nicholas Tierney"
date: "17/05/2020"
output: html_document
---



```{r}
library(tidyverse)
library(geomnet)
library(here)
```

ensure you have installed 'geomnet':

```{r}
# install.packages("remotes")
# library(remotes)
# install_github("sctyner/geomnet")
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Our Turn: Harry Potter characters

There is a connection between two students if one provides emotional support to the other at some point in the book. 
- Code to pull the data together is provided by Sam Tyner [here](https://github.com/sctyner/geomnet/blob/master/README.Rmd#harry-potter-peer-support-network).

```{r read-hp, echo=FALSE}
hp_chars <- read_csv(here::here("data/hp_chars.csv"))
hp_edges <- read_csv(here::here("data/hp_edges.csv"))
```

We need to do a pretty gnarly full join, since we want this for each book:

```{r read-hp, echo=FALSE}
# for each book, expand the edges data and join it to characters
hp_all <- hp_edges %>%
  group_by(book) %>%
  group_map(~ full_join(.x, hp_chars, by = c("name1" = "name"))) %>%
  bind_rows(.id = "book") %>% 
  rename(from_id = name1, 
         to_id = name2)
```

We now have it in that form that we wanted: 

# Harry potter data as nodes and edges

```{r print-hp}
hp_all
```

# Let's plot the characters

```{r plot-hp, echo=TRUE}
ggplot(data = hp_all, 
       aes(from_id = from_id, 
           to_id = to_id)) + 
  geom_net(aes(colour = house, group = house, shape = gender),
           fiteach=T, 
           directed = T, 
           size = 3, 
           linewidth = .5, 
           ealpha = .5, 
           labelon = T, 
           fontsize = 3, 
           repel = T, 
           labelcolour = "black", 
           arrowsize = .5, 
           singletons = FALSE) + 
  scale_colour_manual(values = c("#941B08","#F1F31C", "#071A80", "#154C07")) + 
  facet_wrap(~book, labeller = "label_both", ncol=3) + 
  theme_net() + 
  theme(panel.background = element_rect(colour = 'black'),
        legend.position="bottom")
```

# Your Turn

In the philosphers stone, who has the most connections?

```{r}
hp_long_1 <- hp_edges %>% 
  filter(book == 1) %>% 
  pivot_longer(cols = c("name1", "name2"),
               names_to = "list",
               values_to = "names") 

hp_long_1 %>% 
  count(names, sort = TRUE)
```

In the philosphers stone, who has the least connections?

```{r}
hp_long_1 %>% 
  count(names) %>% 
  arrange(n)
```

Across all the books, can you show the number of connections?

```{r}
# Put it in long form
hp_long <- hp_edges %>% 
  pivot_longer(cols = c("name1", "name2"),
               names_to = "list",
               values_to = "names") 
  
# grouping by book, count up the names 
hp_count <- hp_long %>% 
  group_by(book) %>% 
  count(names)
  
# Can you make a plot of the number of connections across the books?
ggplot(hp_count,
       aes(x = n,
           y = reorder(names, n))) + 
  geom_col()  +
  facet_wrap(~book, scales = "free")
```

