---
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidytext)
library(knitr)
library(lubridate)
opts_chunk$set(echo = TRUE,   
               message = FALSE,
               warning = FALSE,
               collapse = TRUE,
               fig.height = 5,
               fig.width = 8,
               fig.retina = 2,
               fig.align = "center",
               out.width = "100%",
               cache = FALSE)

as_table <- function(...) knitr::kable(..., format='html', digits = 3)
```

---
# Harry Potter books

Using data from Harry potter:

```{r read-harry-potter, echo = FALSE}
# harry potter package from
# https://github.com/bradleyboehmke/harrypotter
hp_books <- read_rds(here::here("data/harry-potter.rds"))
```

# Harry Potter books

Unnest tokens, and use `count` to count up the words within each book:

```{r}
book_words <- hp_books %>% 
  unnest_tokens(word, text) %>% 
  count(book, word, sort = TRUE)
book_words
```

# Inverse document frequency

For the Harry Potter books, we could compute this in a somewhat roundabout as follows:

We can calculate TF IDF using `bind_tf_idf()`

```{r calc-tf-idf}
book_words_counts <- book_words %>%
      anti_join(stopwords_smart) %>% 
      bind_tf_idf(term = word, document = book, n = n)

book_words_counts

# What words were important to the books?

```{r gg-tf-idf, echo=FALSE,message=FALSE}
book_words_counts %>% 
  group_by(book) %>%
  top_n(10, wt = tf_idf) %>%
  ungroup() %>%
  ggplot(aes(x = fct_reorder(word, tf_idf), 
             y = tf_idf, 
             fill = book)) +
  geom_col(show.legend = FALSE) + 
  coord_flip() +
  facet_wrap(~book, scales = "free") +
  scale_y_continuous() +
  theme_minimal() +
  labs(x = NULL, 
       y = "tf-idf")
```

# Your Turn

Explore uncommon / important words in the books?

# Sentiment analysis


```{r print-tidy-ooks}
book_words
```

# Count joyful words in "Chamber of Secrets"

```{r count-joy}
nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")

book_words %>%
  filter(book == "Chamber of Secrets") %>%
  inner_join(nrc_joy) %>%
  arrange(desc(n))
```

# Count joyful words in "Chamber of Secrets"

```{r echo = FALSE}
book_words %>%
  filter(book == "Chamber of Secrets") %>%
  inner_join(nrc_joy) %>%
  arrange(desc(n)) %>% 
  top_n(6, wt = n)
```


"Good" is the most common joyful word, followed by "diary", "found", and "smile". 

These make sense ... except for "diary", and "found", 

--

and ... "white" and "green" ?

# Your turn

- What are the most common "anger" words used in each of the books?
- What are the most common "surprise" words used in each of the books?

# Comparing lexicons

- All of the lexicons have a measure of positive or negative. 
- We can tag the words in Emma by each lexicon, and see if they agree. 

```{r compare-sentiments}
nrc_pn <- get_sentiments("nrc") %>% 
  filter(sentiment %in% c("positive", 
                          "negative"))

secrets_nrc <- book_words %>%
  filter(book == "Chamber of Secrets") %>%
  inner_join(nrc_pn)

secrets_bing <- book_words %>%
  filter(book == "Chamber of Secrets") %>%
  inner_join(get_sentiments("bing")) 

secrets_afinn <- book_words %>%
  filter(book == "Chamber of Secrets") %>%
  inner_join(get_sentiments("afinn"))
```

# Comparing lexicons

```{r show-lexi} 
secrets_nrc
```

# Comparing lexicons

```{r show-lexi-afinn} 
secrets_afinn
```

