---
title: "9a-class-network"
author: "Nicholas Tierney"
date: "17/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(geomnet)
library(janitor)
```

```{r class-names}
class_names <- read_csv(here::here("data/class_names.csv")) %>% clean_names()

class_names
```


```{r class-names}
ggplot(class_names,
       aes(from_id = s1_name, 
           to_id = s2_name)) +
  geom_net(labelon = TRUE)
```


- Try changing the layout method to "circle" or "mds"
- Try setting a random seed, a random number that affects the initial condition for the layout optimisation. You coulde use code like this, substituting in a different number:

```{r}
class_names_subset <- class_names %>%
  filter(team %in% c("angelina", "marina", "sachin"))

class_names_subset
```


- For the class members/tutors, shevin, james, nicole, wai, jamesjing, write out the 5x5 association matrix.

- try exploring the other datasets in the `geomnet` package:

```{r}
library(geomnet)

?geomnet::bikes
?geomnet::blood
?geomnet::email
```


