---
title:
author: 
output: html_document
---

First load up the tidyverse

```{r library}
library(tidyverse)
```

Then read in the data, `rates.csv` and name it `rates`
```{r read-rates}
# ___ <- read_csv("data/___")
rates <- read_csv(here::here("exercises/2a/data/rates.csv"))

head(rates)
```

Now plot the data:

- on the x axis put `date`
- on the y azis put `AUD`

```{r plot-rates}
ggplot(rates, 
       # aes(x = ___, 
       #     y = ___)) + 
       aes(x = date,
           y = AUD)) +
  geom_line()
```

Make a subset of the data that includes the following columns:


- `date`
- `AUD`
- `GBP`
- `JPY`
- `CNY`
- `CAD`

```{r subset-rates}
rates_sub <- rates %>% 
  # select(___, 
  #        ___, 
  #        ___, 
  #        ___, 
  #        ___, 
  #        ___)
  select(date,
         AUD,
         GBP,
         JPY,
         CNY,
         CAD)

rates_sub
```


Now pivot the data into long form, where:

* `names_to` is `currency`
* `values_to` is `rate`

```{r rates-long}
rates_long <- rates_sub %>%
  pivot_longer(cols = -date,
               names_to = "currency", 
               values_to = "rate")
               # names_to = ___, 
               # values_to = ___)
rates_long
```

Now plot the currency rates

* with `date` on the x axis
* with `rate` on the y axis
* And faceted by each currency

```{r plot-all-currencies, fig.width=6, fig.height=10}

ggplot(rates_long, 
       #% aes(x = ___, 
       #%     y = ___)) + 
       aes(x = date,
           y = rate)) +
  geom_line() +
  facet_wrap(~currency, # facetted by currency
             ncol = 1, 
             scales = "free_y")
```

* What do you learn?

Describe the similarities and differences between the currencies. 
