---
title: "more on joins"
author: "Nicholas Tierney"
date: "08/04/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library}
library(tidyverse)
```

```{r read-in}
gap_life <- read_csv(here::here("data/gap_life.csv"))
gap_life_au <- read_csv(here::here("data/gap_life_au.csv"))
gap_income <- read_csv(here::here("data/gap_income.csv"))
gap_income_au <- read_csv(here::here("data/gap_income_au.csv"))
gap_co2 <- read_csv(here::here("data/gap_co2.csv"))
gap_co2_au <- read_csv(here::here("data/gap_co2_au.csv"))

# gap_life <- read_csv(here::here("exercises/4b/data/gap_life.csv"))
# gap_life_au <- read_csv(here::here("exercises/4b/data/gap_life_au.csv"))
# gap_income <- read_csv(here::here("exercises/4b/data/gap_income.csv"))
# gap_income_au <- read_csv(here::here("exercises/4b/data/gap_income_au.csv"))
# gap_co2 <- read_csv(here::here("exercises/4b/data/gap_co2.csv"))
# gap_co2_au <- read_csv(here::here("exercises/4b/data/gap_co2_au.csv"))

```

# Example from lecture

We have information on australian life expectancy and income:

```{r}
gap_life_au
gap_income_au
```

How do we plot life expectancy against income? We need them in the same dataframe

One technique that we've looked at is `bind_cols()`, which binds dataframes together, column wise

```{r}
bind_cols(gap_life_au,
          gap_income_au)
```

But this has problems:

1. It produces messy output (country1, year1)
2. It doesn't work if the data doesn't have the same number of rows

For example, how do we add this co2 data to income or life?

```{r}
gap_co2_au
```

We can't use `bind_cols()`

```{r}
bind_cols(gap_co2_au,
          gap_income_au)
```

We could think about a more complex approach using `filter`, and so on...

But surely this must be a problem that we encounter in data analysis?

Someone must have thought of a solution to this before?

They did! Joins!

We can use ` left_join()` to combine the income and life expectancy data

```{r}
left_join(x = gap_income_au,
          y = gap_life_au,
          by = c("country", "year"))
```

And then what about adding co2 data?

We can add on another join:

```{r}
left_join(x = gap_income_au,
          y = gap_life_au,
          by = c("country", "year")) %>% 
  left_join(gap_co2_au,
            by = c("country", "year"))
```

And we get some missing values for co2, because we don't have c02 values for 2015 and beyond.

So now we can combine that together like so:

```{r}
gap_au <- left_join(x = gap_income_au,
          y = gap_life_au,
          by = c("country", "year")) %>% 
  left_join(gap_co2_au,
            by = c("country", "year"))

gap_au
```

Now we can make a plot!

```{r}
ggplot(gap_au,
       aes(x = gdp,
           y = life_expectancy)) + 
  geom_point()
```

# Next exercise:

Discuss with your partner why these two joins produce different results?

```{r}
left_join(gap_co2_au,
          gap_life_au)

left_join(gap_life_au,
          gap_co2_au)
```

# Next Exercise: What happens when we add data from New Zealand into the mix? Can you produce a dataset that has all data on gdp, life expectancy, and co2, for both Australia and New Zealand, using the joining techniques you've just learnt?

```{r}
gap_life
gap_income
gap_co2
```

