---
title: "Tidy text analysis of Animal Crossing"
author: "Your Name"
date: "10/05/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidytext)
```

# Your turn: Analysing reviews of a video game

This is a continuation of the example we looked at in the lecture:

- User and critic reviews for the game [Animal Crossing](https://www.nintendo.com/games/detail/animal-crossing-new-horizons-switch/) scraped from Metacritc

- This data comes from a [#TidyTuesday challenge](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-05-05/readme.md). 

We can read the data into R directly using the following URLs:

_(Note this requires an internet connection to work)_

```{r}
critics <- readr::read_tsv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-05/critic.tsv')

user_reviews <- readr::read_tsv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-05/user_reviews.tsv')
```

Go through the process of using the critics data to look at the following:

- what are the most used words over the collection of reviews?
- what are the most used words, after removing stop words?
- plot the distrubtion of word frequencies over the collection of reviews
- what is the longest review? what is the shortest review?
- Using the "afinn" add sentiment values to each word in a review
  - for each publication compute the average sentiment for the review
  - are longer reviews more positive?
  - do the grades correlate with the review score, are there any reviews
  with negative sentiments but high scores?
